\documentclass[12pt,oneside]{book}

\input{control.tex}

\begin{document}

\input{title.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN CHAPTERS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\mainmatter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 1/INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\singlespacing
\chapter{Introduction}
\label{ch:intro}
\doublespacing

\pagestyle{fancy}
\pagenumbering{arabic}
%\fancyhead[LE,RO]{\thepage} %puts page number on top left for even pages, top right for odd pages
\chead{}
%\lhead{}
%\rhead{\thepage}
\cfoot{\thepage}
\renewcommand{\headrulewidth}{0pt} %controls thickness of line underneath header- same for footrulewidth
\thispagestyle{fancy}

Earthquakes are like lightning, each strike lighting up the darkness of Earth's interior. Over a century scientists have studied the propagation of the seismic waves, which is the elastic energy released by earthquakes, and used them to explore the deep structures and the dynamic processes of the planet that we all live on. 

Among the different types of seismic waves, surface waves draw the attention of many seismologists as the late-arriving phases which usually have the largest amplitude. They propagate along the shallow part of the earth, and carry information about heterogeneities in crust and upper mantle  that is complementary to that provided by seismic body waves.
Surface waves are dispersed, since different frequencies are sensitive to  structures at different depths, which makes it more challenging to quantitatively describe their propagation. 
They can be measured by different means. One of the most common methods is to estimate phase velocity, the speed at which the phase of each individual frequency varies along the path. 
The frequency-dependent phase velocities can then be converted into the shear velocities at depth through a non-linear inversion.

Based on the particle-motion direction, surface waves can be divided into two categories: Rayleigh waves and Love waves. Although they sample similar parts of the earth, the structures they each suggest  are not always consistent. This so-called Love-Rayleigh discrepancy is an important constraint on the earth's radial anisotropy, which is related to mantle flow pattern, crystal alignment, and melt geometry, among other processes \citep[e.g.][]{Gaherty:2004is,Gaherty:2007cc,Holtzman:2010fta}.

The estimation of surface-wave phase velocity depends on single-station phase measurements. Several techniques have been developed to estimate surface-wave frequency-dependent phase, from utilizing basic Fourier transform \citep{Forsyth:2005id}, to applying multi-channel narrow-band filters \citep[e.g.][]{Levshin:1992ve}, to cross-correlating the waveform with either synthetic waveforms \citep[e.g.][]{Gee:1992ww, Ekstrom:1997ff}, or the waveform from another station on the same great-circle path \citep[e.g.][]{Landisman:1969gt}.

The phase variations at stations can then be inverted for two-dimensional phase-velocity maps. At this stage, several methods are developed under different assumptions. The first category of methods is based on the straight-ray theory, in which the phase observations are represented as integral of phase slowness along the source-receiver path \citep[e.g.][]{Nettles:2008ha}, or between the two stations on the same great-circle path \citep[e.g.][]{Yao:2005ha,foster:2014kna}. The second category attempts to account for ray-bending and finite-frequency effects, which can be substantial for surface waves \citep[e.g.][]{Evernden:1954ui,Zhou:2005fk}. These methods include the application of arrival-angle corrections to two-station measurements \citep{foster:2014kna}, Eikonal tomography \citep{Lin:2009fx}, and finite-frequency kernel estimation \citep{Zhou:2006gna}. The final type of phase-velocity inversion methods utilize amplitude measurements to reduce the multi-pathing interference, including the two-plane-wave method \citep{Forsyth:2005id} and the recently developed Helmholtz tomography \citep{Lin:2011fw}.

Many of these methods are based on the array analysis, which improves our ability to focus on a specific region of interest, regardless of the location of the source. With dense observations, it allows some techniques to increase result resolution by accounting for waveform complexity \citep[e.g.][]{Forsyth:2005id,Lin:2011fw}. However, the array analysis is limited in resolution at long periods due to the finite frequency effect, and it is susceptible to biases associated with limited propagation distances.

The deployment of large, dense arrays has become common in seismic experiments. Substantial amount of data are collected in real time and open to the public. 
Among the experiments, the Transportable Array (TA) under the USArray program stands out by its data volume and coverage area. The TA project includes more than 400 broad-band seismic sensors deployed with $\sim$70-km grid spacing, spanning from the west coast to the east to cover the entire US continent. The communication system installed at the sites allows for real-time data collection and archiving. Such a large and growing dataset is ideal for surface-wave analysis, however an automated system is required to accommodate the speed of data growth.

We develop a new technique to meet the challenge in Chapter~\ref{ch:aswms}. The Automated Surface-Wave phase-velocity Measuring System (ASWMS) is based on the cross-correlation technique to measure frequency-dependent phase delays between all possible nearby station pairs. 
Coherence between the waveforms is used to identify low-quality measurements.  
The measured phase delays, together with single-station amplitude measurements, are then inverted for two-dimensional phase-velocity maps at each frequency.

We successfully automate the entire process. By combining the cross-correlation analysis with an automated data-fetching interface from IRIS DMC, we develop a program that can operate by itself and update the US Rayleigh-wave phase-velocity results on a weekly basis. The program has been adopted by IRIS DMC as a data product (\url{http://ds.iris.edu/ds/products/aswms/}). The phase velocity results,  together with the ASWMS, are available for the public to download through the IRIS website. 

After the methodology development, we turn our focus to its regional application. In Chapter~\ref{ch:png}, we apply the surface-wave analysis in the region near the D'Entrecasteaux Islands (DI), Papua New Guinea (PNG).
The DI are located at the tip of the Woodlark rifting system, where the continental rift is actively transforming into sea-floor spreading \citep[e.g.][]{Taylor:1999ur, Ferris:2006tr}. Previous studies show the crust beneath the DI has been thinned by 30-50\%, while the body-wave tomography results suggest a upwelling mantle \citep{Abers:2002uj, Ferris:2006tr}.
The islands expose high-pressure (HP) terranes comprised of several gneiss domes with diameters of 20-30km. These gneiss domes are mostly felsic in composition (quartzofeldspathic), and one bears the youngest-known ultrahigh-pressure (UHP) coesite eclogite \citep{Baldwin:2004wx, Baldwin:2008gm, Little:2011jy}. The HP/UHP rocks have a composition similar to continental crust. The isotropic analysis reveals an exhumation history at plate-tectonic rates (few cm/year) from $\sim$100-km depth since 5-8 Ma \citep[e.g.][]{Baldwin:2008gm, Gordon:2012hu}.

The exhumation mechanism of the UHP rocks remains debated, with two competing families of models: low-angle extensional unroofing previously subducted UHP rocks, perhaps associated with subduction reversal along a paleo-subduction channel \citep{Hill:1992jd, Webb:2008fc}, and thinning of overlying crust to allow diapiric penetration of buoyant continental rocks to the surface \citep{Ellis:2011jh, Little:2011jy}.

In order to provide better constrains on the geodynamic process in this area, we deployed 31 land-based and 8 ocean bottom broad-band seismometers, with $\sim$20-km station spacing, to cover a $2.5^\circ \times 2.5^\circ$ area around the DI. The stations collected seventeen months of data, and the project team has applied a wide range of observational techniques to these data to image crust and mantle structure in the region.

Here we present the surface-wave analysis on these data, which provides a three-dimensional shear-velocity model of the curst and upper-most mantle. We applied ASWMS to 93 teleseismic earthquakes recorded on the array, and obtained phase-velocity maps across the region within the band between 20-60 s period.
In order to better constrain the shallow structure, we apply the ambient-noise technique to obtain phase velocities at higher frequencies.
The idea of applying cross correlation on ambient noise to recover the green function between two simultaneously recording stations can be traced back as early as \citet{Aki:1957un}. 
It was not until fifty years later that the method became popular because of the availability of dense seismic networks \citep[e.g.][]{Shapiro:2005kz,Bensen:2007hl}. 
In the PNG case, we perform the ambient-noise measurements in the frequency domain, following the algorithm developed \citet{Ekstrom:2009iv} and \citet{Ekstrom:2013dr}. 
A new Bessel-function waveform fitting technique is developed to better estimate the phase delays between the stations. The phase delays are later inverted for phase-velocity maps using a ray-theory tomography method.

The ambient-noise analysis produces high resolution phase-velocity maps at the 10-17~s periods. Together with the earthquake measurements from the ASWMS, we retrieve the robust phase velocities of Rayleigh waves in a wide frequency range (10-60 s in period), which provide good constraint on the crustal and upper mantle structure down to $\sim$100~km.

For each grid on the map, the phase-velocity dispersion curve is then extracted to invert for the shear-velocity structure independently, and the individual 1D models are combined together to produce the three-dimensional model.
This inversion is non-linear and non-unique \citep{Herrmann:2004wl}, and the output can strongly depend on the starting model \citep[e.g.][]{Foti:2009uo}. In order to produce reliable shear-velocity results, we adopt the receiver function results \citep{Abers:2012ul} to constrain the crustal thickness, apply a grid-search technique to find the best initial models, and use a Monte-carlo like procedure to estimate the starting model dependence.
%We adopt the receiver-function results \citep{Abers:2012ul} is adopted to constrain the crustal thickness. We apply a grid-search over various parameters to find the simple three-layer model that best fits the observation. 
%Then the model is randomly perturbed for 100 times, each serves as an initial model to invert for shear-velocity model. 
%The inverted models are then averaged to obtain the final model, and the shear-velocity variance due to the initial-model dependence can be estimated. 
The shear-velocity results show a slow anomaly in the shallow mantle west of Goodenough island, suggesting the existence of un-exhumed continental crust material and/or partial melt.

Although we successfully apply our methods on Rayleigh waves, systematic bias is found in the phase-velocity measurements when  the same techniques are applied on Love waves. 
The bias is mainly generated by overtone interference, which needs to be handled with special care for the Love-wave single-mode measurements \citep[e.g.][]{Thatcher:1969hg,Forsyth:1975tp,Gaherty:1996uf}.

The overtone interference is more acute for Love waves due to the similar group velocities of the fundamental- and higher-mode Love waves, especially for oceanic structures \citep{Nettles:2011bb}. It can produce large phase and amplitude variations for single-branch measurements.  
Such variations do not significantly affect long-path measurements, since the interference pattern is periodical and the bias can be averaged out over the path \citep{Nettles:2011bb}. But they may introduce large bias in the array-based regional studies which involve the estimation of local phase gradient. In the USArray application, \citet{Foster:2014kr} show that the bias due to the interference is profound at middle-band frequencies (60-150~s), even though the array locates in the center of the continent and only shallow events are selected. 
Based on the global and the regional observations, \citet{Nettles:2011bb} and \citet{Foster:2014kr} both suggest the bias is mainly controlled by the interference between the Love-wave fundamental mode and its first overtone.

In Chapter~\ref{ch:overtone} we construct a series of synthetic tests to simulate the pattern of overtone interference and its influence on the single-branch phase-velocity measurements. A correction technique that has the potential to correct this bias is developed and tested on synthetic data.

The target of this dissertation is to better retrieve earth's structure from surface waves, so that we can provide more constraint on geodynamic processes. 
We develop robust and automatic phase-velocity measuring methods for both earthquake and ambient\nobreakdash-noise based data. These techniques are successfully applied on the data from USArray and a regional temporary array at the D'Entrecasteaux Islands, Papua New Guinea. 
Based on the measurement results, we shed light on the effect of thermal and compositional variations on the upper-mantle shear velocities, and the geodynamic process of continental rift extension and UHP rocks exhumation.
We also propose a correction scheme that can reduce the phase-velocity bias due to overtone interference in Love-wave studies, so that mantle radial anisotropy can be better constrained.

\raggedbottom
\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% CHAPTER 2/FIRST PAPER %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\renewcommand\thechapter{\arabic{chapter}}
\singlespacing
\chapter[Surface-Wave Measurement Based on Cross Correlation]{Surface-Wave Phase-Velocity Measurement Based on Multi-Channel Cross Correlation}
\label{ch:aswms}
\doublespacing

\thispagestyle{fancy}

%\footnotesize
\begin{raggedright}
{\bf Note: } A slightly modified version of this chapter has been submitted to Geophysical Journal International (2014).
\footnote{AUTHORS:  Ge Jin$^a$*,  James Gaherty$^a$\\
$^a$ Lamont-Doherty Earth Observatory, Columbia University, 61 Route 9W, Palisades, NY 10964, USA\\
* corresponding author: ge.jin@ldeo.columbia.edu}
\end{raggedright}
%\linenumbers
\normalsize

\input{aswms.tex}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 3 PNG
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\singlespacing
\chapter[Shear velocity structure of the DI, PNG]{Shear Velocity Structure of the D'Entrecasteaux Islands, Papua New Guinea from Rayleigh Wave Tomography}
\label{ch:png}
\doublespacing

\thispagestyle{fancy}
\begin{raggedright}
%\footnotesize
\footnote{AUTHORS:  Ge Jin$^a$*,  James Gaherty$^a$, Geoff Abers$^b$, Younghee Kim$^c$, Zachary Eilon$^a$, Roger Buck$^a$\\
$^a$ Lamont-Doherty Earth Observatory, Columbia University, 61 Route 9W, Palisades, NY 10964, USA\\
$^b$ Department of Earth and Atmospheric Science, Cornell University\\
$^c$ School of Earth and Environmental Sciences, Seoul National University.\\
* corresponding author: ge.jin@ldeo.columbia.edu}
\end{raggedright}
%\linenumbers
\normalsize

\input{png.tex}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 4 overtone %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\singlespacing
\chapter[Love-wave Overtone Interference]{Love-Wave Phase-velocity Estimation in the Presence of Multi-mode Interference}
\label{ch:overtone}
\doublespacing

\thispagestyle{fancy}
\begin{raggedright}
%\footnotesize
\footnote{AUTHORS:  Ge Jin$^a$*,  James Gaherty$^a$, Anna Foster$^b$, Pei-ying Lin$^a$\\
$^a$ Lamont-Doherty Earth Observatory, Columbia University, 61 Route 9W, Palisades, NY 10964, USA\\
$^b$ Earth Observatory of Singapore\\
* corresponding author: ge.jin@ldeo.columbia.edu}
\end{raggedright}
%\linenumbers
\normalsize

\input{overtone.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Chapter 5 conclusion %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\singlespacing
\chapter[Conclusion]{Concluding Remarks}
\label{ch:conclusion}
\doublespacing

In this dissertation, we develop several new surface-wave measuring methods and apply them on different datasets to understand local tectonic structures.

In Chapter~\ref{ch:aswms}, we develop the Automated Surface-Wave Measuring System (ASWMS), which can retrieve surface-wave phase velocities from dense-array data without human intervention. We produce robust Rayleigh-wave phase-velocity maps of the US continent at the periods of 20-100s using the USArray data, and work with IRIS to provide weekly-updated results for the scientific community and the general public. The phase-velocity maps clearly outline the major geological structures in the US continent. The spectral analysis of these maps implies that in the upper mantle, thermal variation affects shear velocity more than compositional variations.

In Chapter~\ref{ch:png}, we apply surface-wave analysis in the region near the D'Entrecasteaux Islands, Papua New Guinea. We retrieve phase-velocity variations using both ambient-noise and earthquake signals. The phase velocities in a broad period range (10-60 s) are inverted for the shear-velocity structure in the crust and the upper mantle. The results imply the existence of un-exhumed crustal material and/or partial melt in the shallow mantle west of the most recently exhumed UHP rocks on Goodenough Island.

In Chapter~\ref{ch:overtone}, we focus on the effect of Love-wave overtone interference on the array-based phase-velocity measurements. We simulate the interference pattern by two plane waves travelling together at different phase velocities. We also develop a correction scheme that can significantly reduce the phase-velocity bias and variance due to the interference. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\backmatter
\cleardoublepage
\normalsize
\singlespacing
\bibliographystyle{elsarticle-harv}
\addcontentsline{toc}{chapter}{References}
\bibliography{thesis_ref}
\clearpage

\end{document}  
